Simplifying Lotka--Volterra interaction matrices by species grouping: optimal groupings are unpredictable

### 2022-05-17 article.rnw
Mention the comparison of groupings to species traits in the introduction.
Set echo=FALSE as a global default for knitr chunks in the article.

<<include=FALSE>>=
	knitr::opts_chunk$set(results = 'asis') #Text is grey without this.
	knit_hooks$set(inline = function(x) {x}) #Disable automatic formatting of in-line material.
@

@@

We propose that species might be best partitioned into groups without regard to any \emph{a priori} definition but instead based purely upon which grouping gives the most parsimonious model.
% DBS: We can say something here about how this data-informed approach has worked very well for other clustering problems, such as modularity in networks. I can point to some very good references along those lines.
In this study, we investigate the usefulness and correctness of such an approach using experimental data from competition between different \emph{Drosophila} species provided by \textcite{terry2021natural}.
We fit a Lotka--Volterra model to these data to find the groupings that give the most parsimonious model.
While doing so, we group species independently with regard to their effects upon other species and their responses to other species, to find whether the data-informed groupings differ between these two cases.
%TODO: Mention the comparison to traits (body size)?
% DBS: If we leave in the traits, then we should mention them at least briefly.

@@

\begin{figure}
	<<echo=FALSE>>=
		hist(brute_data$statistics[[ic]],
			main="",
			xlab=ic_name,
			xaxs='i',
			yaxs='i'
		)
		abline(v = brute_data$fully_separated_statistics[[ic]], lwd=2, lty="dashed", col="green")
		abline(v = brute_data$fully_grouped_statistics[[ic]], lwd=2, lty="dashed", col="red")
	@

@@

	<<echo=FALSE>>=
		best_model_index <- which.max(brute_data$statistics[[ic_weight]])
		best_row_grouping <- brute_data$get_row_grouping(best_model_index)
		best_col_grouping <- brute_data$get_col_grouping(best_model_index)
		best_alpha_matrix <- brute_data$parameters[[best_model_index]]$alpha_values

@@

	\begin{subfigure}{\textwidth}
		\centering
		<<echo=FALSE>>=
			grouped_alpha_matrix(
				best_species_names,
				best_row_grouping,
				best_col_grouping,
				best_alpha_matrix
			)
		@
		\caption{The grouped alpha matrix.}
		\label{fig:best-alpha-matrix}
	\end{subfigure}

@@

	\begin{subfigure}{0.49\textwidth}
		\centering
		<<echo=FALSE>>=
			minimal_coclassification_kable(
				best_col_coclassification,
				colourmap="coclassification", diagonal_colour="coclassification diagonal"
			)
		@
		\caption{The coclassification matrix of rows.}
		\label{fig:best-coclassification-cols}
	\end{subfigure}
	%
	\begin{subfigure}{0.49\textwidth}
		\centering
		<<echo=FALSE>>=
			minimal_coclassification_kable(
				best_row_coclassification,
				colourmap="coclassification", diagonal_colour="coclassification diagonal"
			)
		@
		\caption{The coclassification matrix of columns.}
		\label{fig:best-coclassification-rows}
	\end{subfigure}

@@

% DBS: Most journals frown upon or do not allow overly verbose subtitles on figures. So I've cut them down to be a bit more concise.
% DBS: But the fact that the labels say effects and responses reminds me of a lingering "doubt" I had when reading the main text. We talk almost entirely about rows and columns but there are a number of junctures where I feel like it would be easier to follow for people who aren't us if we were talking about responses and effects. As long as we're consistent, I'm fine either way. But we need to remember we're writing the paper for the audience and not for ourselves. ;)
\begin{figure}
	\centering
	\begin{subfigure}{\textwidth}
		\centering
		<<echo=FALSE>>=
			weighted_coclassification_kable(
				weighted_col_coclassification_matrix[species_output_order,species_output_order],
				colourmap="coclassification", diagonal_colour="coclassification diagonal"
			)
		@
		\caption{Coclassification of competitive effects}
		\label{fig:coclassification-cols}
	\end{subfigure}

@@

	\begin{subfigure}{\textwidth}
		\centering
		<<echo=FALSE>>=
			weighted_coclassification_kable(
				weighted_row_coclassification_matrix[species_output_order,species_output_order],
				colourmap="coclassification", diagonal_colour="coclassification diagonal"
			)
		@
		\caption{Coclassification of competitive responses}
		\label{fig:coclassification-rows}
	\end{subfigure}

>>>>

<<include=FALSE>>=
	knitr::opts_chunk$set(echo = FALSE) #Never show the code, only output.
	knitr::opts_chunk$set(results = 'asis') #Text is grey without this.
	knit_hooks$set(inline = function(x) {x}) #Disable automatic formatting of in-line material.
@

@@

We propose that species might be best partitioned into groups without regard to any \emph{a priori} definition but instead based purely upon which grouping gives the most parsimonious model.
% DBS: We can say something here about how this data-informed approach has worked very well for other clustering problems, such as modularity in networks. I can point to some very good references along those lines.
In this study, we investigate the usefulness and correctness of such an approach using experimental data from competition between different \emph{Drosophila} species provided by \textcite{terry2021natural}.
We fit a Lotka--Volterra model to these data to find the groupings that give the most parsimonious model, and compare these groupings to similarity based upon a species trait (body size).
While doing so, we group species independently with regard to their effects upon other species and their responses to other species, to find whether the data-informed groupings differ between these two cases.

@@

\begin{figure}
	<<>>=
		hist(brute_data$statistics[[ic]],
			main="",
			xlab=ic_name,
			xaxs='i',
			yaxs='i'
		)
		abline(v = brute_data$fully_separated_statistics[[ic]], lwd=2, lty="dashed", col="green")
		abline(v = brute_data$fully_grouped_statistics[[ic]], lwd=2, lty="dashed", col="red")
	@

@@

	<<>>=
		best_model_index <- which.max(brute_data$statistics[[ic_weight]])
		best_row_grouping <- brute_data$get_row_grouping(best_model_index)
		best_col_grouping <- brute_data$get_col_grouping(best_model_index)
		best_alpha_matrix <- brute_data$parameters[[best_model_index]]$alpha_values

@@

	\begin{subfigure}{\textwidth}
		\centering
		<<>>=
			grouped_alpha_matrix(
				best_species_names,
				best_row_grouping,
				best_col_grouping,
				best_alpha_matrix
			)
		@
		\caption{The grouped alpha matrix.}
		\label{fig:best-alpha-matrix}
	\end{subfigure}

@@

	\begin{subfigure}{0.49\textwidth}
		\centering
		<<>>=
			minimal_coclassification_kable(
				best_col_coclassification,
				colourmap="coclassification", diagonal_colour="coclassification diagonal"
			)
		@
		\caption{The coclassification matrix of rows.}
		\label{fig:best-coclassification-cols}
	\end{subfigure}
	%
	\begin{subfigure}{0.49\textwidth}
		\centering
		<<>>=
			minimal_coclassification_kable(
				best_row_coclassification,
				colourmap="coclassification", diagonal_colour="coclassification diagonal"
			)
		@
		\caption{The coclassification matrix of columns.}
		\label{fig:best-coclassification-rows}
	\end{subfigure}

@@

% DBS: Most journals frown upon or do not allow overly verbose subtitles on figures. So I've cut them down to be a bit more concise.
% DBS: But the fact that the labels say effects and responses reminds me of a lingering "doubt" I had when reading the main text. We talk almost entirely about rows and columns but there are a number of junctures where I feel like it would be easier to follow for people who aren't us if we were talking about responses and effects. As long as we're consistent, I'm fine either way. But we need to remember we're writing the paper for the audience and not for ourselves. ;)
\begin{figure}
	\centering
	\begin{subfigure}{\textwidth}
		\centering
		<<>>=
			weighted_coclassification_kable(
				weighted_col_coclassification_matrix[species_output_order,species_output_order],
				colourmap="coclassification", diagonal_colour="coclassification diagonal"
			)
		@
		\caption{Coclassification of competitive effects}
		\label{fig:coclassification-cols}
	\end{subfigure}

@@

	\begin{subfigure}{\textwidth}
		\centering
		<<>>=
			weighted_coclassification_kable(
				weighted_row_coclassification_matrix[species_output_order,species_output_order],
				colourmap="coclassification", diagonal_colour="coclassification diagonal"
			)
		@
		\caption{Coclassification of competitive responses}
		\label{fig:coclassification-rows}
	\end{subfigure}
